
Title: "Reproducible Research Project 1"
==========================================
The data for this assignment can be downloaded from the course web site:

Dataset: [Activity monitoring data [52K]](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

The variables included in this dataset are:

* steps: Number of steps taking in a 5-minute interval (missing values are coded as \color{red}{\verb|NA|}NA)
* date: The date on which the measurement was taken in YYYY-MM-DD format
* interval: Identifier for the 5-minute interval in which measurement was taken

Set global options - show r code 
```{r setoptions, echo=TRUE}
library(ggplot2)
library(knitr)
opts_chunk$set(echo=TRUE)
```

#1. Code for reading in the dataset and/or processing the data
``` {r}
activity<-read.csv("./activity/activity.csv")
head(activity)
str(activity)
summary(activity)
```
Process/transform the data (if necessary) into a format suitable for your analysis
```{r convertdate}
#convert date from factor to date format
activity$date<-as.Date(activity$date, format = "%Y-%m-%d")
head(activity)
str(activity)
```
#2. Histogram of the total number of steps taken each day
```{r totalstepsdaily}
total.steps.perday<-with(activity,tapply(steps, date, sum, na.rm=T))
str(total.steps.perday)
hist(total.steps.perday)
mean.daily.steps<-as.integer(mean(total.steps.perday))
median.daily.steps<-median(total.steps.perday)
print(mean.daily.steps)
print(median.daily.steps)
```
#3. Mean and median number of steps taken each day
```{r}
mean.steps.perday<-as.integer(mean(total.steps.perday, na.rm=T))
print(mean.steps.perday)
```
```{r}
median.steps<-as.integer(median(total.steps.perday, na.rm=T))
print(median.steps)
```
#4. Time series plot of the average number of steps taken
Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
library(dplyr,quietly = TRUE)
library(ggplot2)
avg.steps<- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
avg.steps.table<-data.frame(interval=as.numeric(names(avg.steps)), avg.steps=as.integer(avg.steps))
p <- ggplot(avg.steps.table, aes(x=interval, y=avg.steps), xlab = "Interval", ylab="Average Number of Steps")
p + geom_line(color="blue")+xlab("Interval")+ylab("Average Number of Steps")+ggtitle("Average Steps Per Interval")

```
#5. The 5-minute interval that, on average, contains the maximum number of steps
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max.steps<-max(avg.steps.table$avg.steps)
max.interval<-avg.steps.table[avg.steps.table$avg.steps==max.steps, 1]
print(max.interval)
```
#6. Code to describe and show a strategy for imputing missing data
first, check the amount of missing values
```{r}
activity<-read.csv("./activity/activity.csv")
missing.steps<-sum(is.na(activity$steps))
print(missing.steps)
missing.pct<-mean(is.na(activity$steps))
print(missing.pct)
```
Strategy: use mean step per interval of that day for imputation
```{r}
library(dplyr)
activity<-read.csv("./activity/activity.csv")
activity$date<-as.Date(activity$date, format = "%Y-%m-%d")
```

Create data frame of mean interval steps by date
```{r}
meandf <- activity %>%
  group_by(date) %>%
  summarise_at(vars(-interval), funs(mean(., na.rm=TRUE)))
```

There are 8 days missing steps data. Use mean daily interval steps (37) to replace
```{r}
meandf$steps[which(is.na(meandf$steps))]<-mean(meandf$steps, na.rm=TRUE)
meandf$steps<-as.integer(meandf$steps)
names(meandf) <-c("date", "meansteps")
meandf
```

Merge activity with mean steps information, and replace missing values with mean steps per interval for that date
```{r}
activity2<-merge(activity, meandf, by="date", all.x=TRUE)
activity2$steps[which(is.na(activity$steps))]<-activity2$meansteps[which(is.na(activity$steps))]
```
#7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
total.steps.perday2<-with(activity2,tapply(steps, date, sum, na.rm=T))
hist(total.steps.perday2)
```
```{r}
mean.daily.steps<-as.integer(mean(total.steps.perday2))
median.daily.steps<-median(total.steps.perday2)
print(mean.daily.steps)
print(median.daily.steps)
```

original data (with NA), mean daily step: 10766               median daily steps: 10765
After imputation:        mean daily step: 10751               median daily steps: 10656

#8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
library(lubridate)
activity2$daytype <- ifelse(wday(activity2$date, label=TRUE) == "Sat" | wday(activity2$date,label=TRUE) == "Sun", "weekend", "weekday")

#average steps per interval for weekdays
avgsteps.weekday <- tapply(activity2[activity2$daytype == "weekday",]$steps, activity2[activity2$daytype == "weekday" ,]$interval, mean, na.rm = TRUE)

#average steps per interval for weekends
avgsteps.weekend <- tapply(activity2[activity2$daytype == "weekend" ,]$steps, activity2[activity2$daytype == "weekend" ,]$interval, mean, na.rm = TRUE)

intervals<-as.numeric(names(avgsteps.weekday ))
stepwkday<-data.frame(Interval=as.numeric(intervals), Steps=as.integer(avgsteps.weekday), day=rep("weekday", length(intervals)))
stepwkend<-data.frame(Interval=as.numeric(intervals), Steps=as.integer(avgsteps.weekend), day=rep("weekend", length(intervals)))
steps.df<-rbind(stepwkday, stepwkend)

qplot(Interval, Steps, data=steps.df, facets=day~., geom=c("line"), color=day)
```

